---
title: "CK119 late healthy organoid: initial clustering"
author: "Javier Perales-Patón - javier.perales@bioquant.uni-heidelberg.de - ORCID: 0000-0003-0780-6683" 
license: "GPL-v3"
output: github_document
---

```{r setup, include=FALSE}
Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio/bin/pandoc")
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and auxiliar functions

```{r libs}
set.seed(1234)
suppressPackageStartupMessages(require(Seurat))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(clustree))
suppressPackageStartupMessages(require(cowplot))
source("../src/seurat_fx.R")
```

## Read data into a Seurat Object
```{r data, cache=FALSE}
SeuratObject <- getSeuratObject(path = "../data/sc/CK119_organoid/filtered_feature_bc_matrix/",
                     project_name = "CK119_organoid", 
                     mt.pattern = "^MT-", min.cells = 5, 
                     min.features = 200)
```


## Define output directory
```{r def_out}
# Define output directory
OUTDIR <- paste0("./output/1_initial_clustering/")
if(! dir.exists(OUTDIR)) dir.create(OUTDIR, recursive = TRUE)
```

## Quality control

```{r QC,fig.width=14, fig.height=4}
print(VlnPlot(SeuratObject, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3))

print(CombinePlots(plots = list(FeatureScatter(SeuratObject, 
					       feature1 = "nCount_RNA", 
					       feature2 = "percent.mt"),
                                FeatureScatter(SeuratObject, 
					       feature1 = "nCount_RNA", 
					       feature2 = "nFeature_RNA")))
)
```

## Data cleasing: gene and cell filtering
To avoid unnecessary sparsity and noise in downstream analysis,    
* We discard cells with low (<200) and high (>6000) number of genes detected to 
avoid bad quality and doublets, repectively.    
* We discard genes not detected across cells.    
For this we use cutoffs in concordance with kidney tissue, where Proximal tubule
cells might have a high content of mitochondrial genes (<20%). Later we will
apply diagnostics to see if this is related.

```{r cars}
nSamples_before <- ncol(SeuratObject)
SeuratObject <- subset(SeuratObject, nFeature_RNA > 200 & nFeature_RNA < 6000 & percent.mt < 20)
nSamples_after <- ncol(SeuratObject)
```

Doing so, `r nSamples_after` out of an initial `r nSamples_before` total were retrieve
for the analysis, thus discarding a `r (nSamples_before - nSamples_after)` cells.

## Pre-processing the data for cell clustering and cell-type assignment
First, we normalize the data using default parameters
```{r norm, cache=FALSE}
SeuratObject <- NormalizeData(SeuratObject)
```

Second, we perform a feature selection based on `vst` method to select 
high variable genes (top 2000) for cell clustering.
```{r sel, cache=FALSE}
SeuratObject <- FindVariableFeatures(SeuratObject, selection.method = "vst", nfeatures = 2000)

print(LabelPoints(VariableFeaturePlot(SeuratObject),
                  points=VariableFeatures(SeuratObject)[1:10],repel = TRUE))
```

Third, we center and scale the data prior PCA:

```{r scale,cache=FALSE}
SeuratObject <- Seurat_scaledata(SeuratObject)
```

Finally we perform PCA, selecting a number of PCs that are relevant based on the
elbow plot. We tested that our results are not sensitive to this parameter, and
the arbitrary selection of number of PCs do not change final conclusions.
```{r pca, cache=FALSE}
SeuratObject <- RunPCA(SeuratObject, features = VariableFeatures(SeuratObject), npcs = 50)
print(ElbowPlot(SeuratObject,ndims = 50) + geom_vline(xintercept = 25, col="red"))
```

We decided to use 25 PCs for the clustering.

We also calculate a dissociation score
```{r dissociation}
dissociation <- read.csv("../data/Prior/scrnaseq-digestion-paper_coregene_df-FALSE-v3.csv",
			 stringsAsFactors=FALSE)
dissTop40 <- intersect(head(dissociation$gene_symbol, 40), rownames(SeuratObject))
SeuratObject <- AddModuleScore(SeuratObject, features = list("dissociation"=dissTop40))
colnames(SeuratObject@meta.data)[which(colnames(SeuratObject@meta.data)=="Cluster1")] <- "Dissociation"
```

And cellcycle phase
```{r CellCycle}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes
SeuratObject <- CellCycleScoring(SeuratObject,
				 s.features = s.genes,
				 g2m.features = g2m.genes)
```



## Cell clustering
We are going to use Shared-Nearest Neighbour with Graph partitioning for cell clustering.
We run the algorithm with multiple resolutions. Later this is read-out to give 
an idea how these are consistent from a lower to high number of expected cell populations
in the sample.   

```{r clust, cache=FALSE}
SeuratObject <- FindNeighbors(SeuratObject, dims = 1:25)
SeuratObject <- FindClusters(SeuratObject, resolution = seq(from=0.1, to=1, by=0.1))

# We redefine the final partitioning with resolution 0.5
SeuratObject$seurat_clusters <- SeuratObject$RNA_snn_res.0.5 
Idents(SeuratObject) <- SeuratObject$RNA_snn_res.0.5 
```

## Diagnostics of unsupervised clustering

We will investigate how the selection of multiple resolutions affects the partition
into individual cell clusters.
```{r clustree, fig.width=7, fig.height=10}
clustree(SeuratObject, prefix = "RNA_snn_res.")
```

We also checked whether mithochondrial expression drives the clustering
```{r clustree_mt, fig.width=7, fig.height=10}
clustree(SeuratObject, prefix = "RNA_snn_res.", 
	 node_colour="percent.mt", node_colour_aggr="mean")
```

```{r clustree_nFeat, fig.width=7, fig.height=10}
clustree(SeuratObject, prefix = "RNA_snn_res.", 
	 node_colour="nFeature_RNA", node_colour_aggr="median")
```

```{r vln_badQual_mtNfeat}
VlnPlot(SeuratObject, 
	feature=c("percent.mt", "nFeature_RNA", 
		"Dissociation", "G2M.Score"), 
	pt.size=0.4,
	ncol=2)
```

## Clustering decision

We decided to chose resolution 0.5 for the initial clustering. We will save this 
output to archive the outcome. Herein, first column is the cell 
barcode, and rest of columns are clusters where each cell belong to across 
multiple resolutions. The ones that are active for the study are those related to
resolution 0.5 (`RNA_snn_res.0.5` and `seurat_clusters`).

```{r clustering_record}
# 1 Clustering outcome
saveClusteringOutcome(SeuratObject , assay="RNA", fl=paste0(OUTDIR,"/init_clustering.tsv"))
# 2 Initial idents (same as seurat_clusters)
saveActiveIdents(SeuratObject, fl=paste0(OUTDIR,"/active_idents.tsv"))
```

## Non-linear dim reduction (umap)
```{r umap, cache=FALSE}
SeuratObject <- RunUMAP(SeuratObject, dims = 1:25)
DimPlot(SeuratObject, reduction="umap") + ggtitle(Project(SeuratObject))
```

## Archive processed data for downstream analysis
```{r, mkdir_data}
DATA_DIR <- paste0(OUTDIR,"/data")
if(!dir.exists(DATA_DIR)) dir.create(DATA_DIR)
```

```{r save_S}
saveRDS(SeuratObject, paste0(DATA_DIR,"/SeuratObject.rds"))
```                                                                                                                                                                                                         

## Session info

```{r}
sessionInfo()
{                                                                                                                                                                                                           
sink(file=paste0(OUTDIR,"/sessionInfo.txt"))
print(sessionInfo())
sink()
}
```
