---
title: "Control2 whole tissue biopsy, healthy: cell assignment"
author: "Javier Perales-Paton - javier.perales@bioquant.uni-heidelberg.de"
license: "GPL-v3"
output: github_document
---


```{r setup, include=FALSE}
Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio/bin/pandoc")
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and auxiliar functions
```{r libs}
set.seed(1234)
suppressPackageStartupMessages(require(Seurat))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(GSEABase))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(genesorteR))
suppressPackageStartupMessages(require(ComplexHeatmap))
suppressPackageStartupMessages(require(clustree))
suppressPackageStartupMessages(require(cowplot))
suppressPackageStartupMessages(require(openxlsx))
source("../src/seurat_fx.R")
```

## Load SeuratObject with initial clustering outcome
```{r load_init}
SeuratObject <- readRDS("./output/1_initial_clustering/data/SeuratObject.rds")
```

## Define output directory
```{r def_out}
# Define output directory
OUTDIR <- paste0("./output/2_cell_assignment/")
if(! dir.exists(OUTDIR)) dir.create(OUTDIR, recursive = TRUE)
```

## Cell marker extraction
We are going to use two methods for the extraction of cell markers.   
1. Differential gene expression using wilcox test from Seurat. Only positive
   and consistent markers will be tested (with default parameters).   
2. GenesorteR to get posterior probabilities for each gene of observing a certain
cell population.

### 1 wilcox test

```{r wilcox, cache=FALSE}
up <- setNames(vector("list",length=length(levels(SeuratObject))), 
               levels(SeuratObject))
for(idx in names(up)) {
  up.idx <- FindMarkers(SeuratObject,ident.1 = idx, 
                        ident.2 = setdiff(levels(SeuratObject), idx), only.pos=T)
  cols_names <- colnames(up.idx)
  
  # Add two extra cols
  up.idx$cluster <- idx
  up.idx$gene <- rownames(up.idx)
  
  up[[idx]] <- up.idx
}
```

### 2 Gene sorter

```{r genesort, cache=FALSE}
sg <- sortGenes(SeuratObject@assays$RNA@data, Idents(SeuratObject))
```

Finally we save the markers for manual exploration:

```{r init_markers_save, warning=FALSE}
saveMarkers.Excel(up, sg, SeuratObject, OUTDIR)
saveMarkers.CSV(up, OUTDIR)
saveMarkers.CSV(sg, OUTDIR)
```

## Cell assignment

```{r cellQual}
FeaturePlot(SeuratObject, 
	    features=c("percent.mt", "nFeature_RNA", 
		       "S.Score", "G2M.Score"), 
	    label=TRUE, label.size=7)
```

```{r CellCycle, fig.width=7, fig.height=5}
VlnPlot(SeuratObject, features=c("S.Score", "G2M.Score"))
```

We will transfer label from Lake et al 2019 to corroborate this
```{r Lake2019_transfer}
Lake <- readRDS("../data/Lake2019/01_Seurat/data/S.rds")
Lake <- NormalizeData(Lake, verbose=FALSE)
Lake <- FindVariableFeatures(Lake, selection.method="vst", nfeatures=2000, verbose=FALSE)
anchors <- FindTransferAnchors(reference = Lake, query = SeuratObject)
predictions <- TransferData(anchorset = anchors, refdata = Lake$cluster_abbn)

res <- table(Idents(SeuratObject), predictions$predicted.id)
res2 <- sapply(rownames(res), function(idx) sort(res[idx,], decreasing=TRUE)[1:5], simplify=FALSE)
```

```{r rename_idents}
ren_id <- unlist(lapply(res2, function(z) names(z)[1]))
ren_id <- gsub("(-|_)_?[0-9]+", "", ren_id)
# Correction by hand
ren_id[names(ren_id)=="7"] <- "iPT"
ren_id[names(ren_id)=="14"] <- "PEC"
# Index them
tag_dups <- ren_id[duplicated(ren_id)]
for(tag in tag_dups) {
	idx <- which(ren_id==tag)
	ren_id[idx] <- paste0(tag,"_",1:length(idx))
}

print(ren_id)
stopifnot(length(ren_id)==length(levels(SeuratObject)))
```

```{r cell_ann}
stopifnot(all(names(ren_id) %in% levels(SeuratObject)))
stopifnot(length(ren_id) == length(levels(SeuratObject)))

SeuratObject <- RenameIdents(SeuratObject, ren_id)
SeuratObject$init_assign <- factor(as.character(Idents(SeuratObject)))

colnames(sg$condGeneProb) <- ren_id[colnames(sg$condGeneProb)]
colnames(sg$specScore) <- ren_id[colnames(sg$specScore)]
```

## Kidney markers

```{r load_markers}
AdultOrganoid <- getGmt("../data/Prior/adultorganoid_markers.gmt")
AdultOrganoid <- GeneSetCollection(lapply(AdultOrganoid, function(GS) {
			  GS@geneIds <- intersect(GS@geneIds, rownames(SeuratObject))
			  GS@geneIds <- intersect(GS@geneIds, rownames(sg$condGeneProb))
			  return(GS)
	    }))
AdultOrganoid <- AdultOrganoid[unlist(lapply(geneIds(AdultOrganoid), length)) > 0]
```

```{r AdultOrganoid_mhp, fig.width=16, fig.height=14, dpi=300, warning=FALSE}
DoMultiHeatmap(SeuratObject, sg, GSC = AdultOrganoid, assay = "RNA", show_plot = TRUE)
```

```{r AdultOrganoid_dplot, fig.width=16, fig.height=14, dpi=300, warning=FALSE}
DoMultiDotPlot(SeuratObject, sg, GSC = AdultOrganoid)
```

```{r ClustReport, fig.width=16, fig.height=17, dpi=300, warning=FALSE}
DoClustReport(SeuratObject, sg, AdultOrganoid, show_NCells=TRUE)
```

## Cyst markers

```{r load_markers2}
Cyst <- getGmt("../data/Prior/cyst.gmt")
Cyst <- GeneSetCollection(lapply(Cyst, function(GS) {
			  GS@geneIds <- intersect(GS@geneIds, rownames(SeuratObject))
			  GS@geneIds <- intersect(GS@geneIds, rownames(sg$condGeneProb))
			  return(GS)
	    }))
Cyst <- Cyst[unlist(lapply(geneIds(Cyst), length)) > 0]
```

```{r Cyst_mhp, fig.width=16, fig.height=6, dpi=300, warning=FALSE}
DoMultiHeatmap(SeuratObject, sg, GSC = Cyst, assay = "RNA", show_plot = TRUE)
```

```{r Cyst_dplot, fig.width=16, fig.height=6, dpi=300, warning=FALSE}
DoMultiDotPlot(SeuratObject, sg, GSC = Cyst)
```

## UMAP plot
```{r}
DimPlot(SeuratObject, reduction = "umap", label=TRUE)
```

## Archive processed data for downstream analysis

```{r assign_record}
# new idents
write.table(data.frame("Ident"=SeuratObject@active.ident),
            file=paste0(OUTDIR,"/active_idents.tsv"),
            sep="\t", col.names = NA, row.names = TRUE, quote=TRUE)
```

```{r, mkdir_data}
DATA_DIR <- paste0(OUTDIR,"/data")
if(!dir.exists(DATA_DIR)) dir.create(DATA_DIR)
```

```{r save_S}
saveRDS(SeuratObject, paste0(DATA_DIR,"/SeuratObject.rds"))
saveRDS(up, file=paste0(OUTDIR,"/wilcox_up.rds"))
saveRDS(sg, file=paste0(OUTDIR,"/genesorter_out.rds"))

# After Rename Idents
saveMarkers.Excel(up, sg, SeuratObject, OUTDIR)

```                                                                                                                                                                                                         

## Session info

```{r}
sessionInfo()
{                                                                                                                                                                                                           
sink(file=paste0(OUTDIR,"/sessionInfo.txt"))
print(sessionInfo())
sink()
}
```
