---
title: "ADPKD marker differential expression"
author: "Javier Perales-Paton - javier.perales@bioquant.uni-heidelberg.de"
license: "GPL-v3"
output: github_document
---


```{r setup, include=FALSE}
Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio/bin/pandoc")
knitr::opts_chunk$set(echo = TRUE)

# Output directory
OUTDIR <- "./PKD_vs_Healthy/"
if(!dir.exists(OUTDIR)) dir.create(OUTDIR);
# Figures
FIGDIR <- paste0(OUTDIR, "/figures/")
knitr::opts_chunk$set(fig.path=FIGDIR)
knitr::opts_chunk$set(dev=c('png', 'pdf'))
knitr::opts_chunk$set(dpi=300)
# Data
DATADIR <- paste0(OUTDIR, "/data/")
if(!dir.exists(DATADIR)) dir.create(DATADIR);
```


## Load libraries and auxiliar functions

```{r libs}
set.seed(1234)
suppressPackageStartupMessages(require(Seurat))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(clustree))
suppressPackageStartupMessages(require(cowplot))
suppressPackageStartupMessages(require(GSEABase))
suppressPackageStartupMessages(require(clusterProfiler))
suppressPackageStartupMessages(require(openxlsx))
suppressPackageStartupMessages(require(ComplexHeatmap))
source("../src/seurat_fx.R")
```

## Load SeuratObject with initial clustering outcome
```{r load_init}
late <- readRDS(paste0("../Individual_analysis_CK119_late_organoid/",
			"output/2_cell_assignment/data/SeuratObject.rds")
)
late$final_Ident <- Idents(late)
late$orig.ident <- factor("Healthy")

ADPKD_PDK1KO <- readRDS(paste0("../Individual_analysis_JX2_PKD1KO_organoid/",
		     	"output/2_cell_assignment/data/SeuratObject.rds"))
ADPKD_PDK1KO$final_Ident <- Idents(ADPKD_PDK1KO)
ADPKD_PDK1KO$orig.ident <- factor("PDK1KO")

ADPKD_PDK2KO <- readRDS(paste0("..//Individual_analysis_JX3_PKD2KO_organoid/",
		     	"output/2_cell_assignment/data/SeuratObject.rds"))
ADPKD_PDK2KO$final_Ident <- Idents(ADPKD_PDK2KO)
ADPKD_PDK2KO$orig.ident <- factor("PDK2KO")
```

## Merging
```{r merge}
S <- merge(late, list(ADPKD_PDK1KO, ADPKD_PDK2KO))
```

## Refining data

Check unique cell types among organoids
```{r}
unique(gsub("_[0-9]", "", levels(S)))
```

Discard low quality clusters and proliferating
```{r discard}
S <- S[, !grepl("(Prolif|\\:Low|\\:High)", Idents(S), ignore.case = TRUE)]
Idents(S) <- droplevels(Idents(S))
unique(gsub("_[0-9]", "", levels(S)))
```

Pooling cell types
```{r}
Idents(S) <- factor(gsub("_[0-9]", "", as.character(Idents((S)))))
```

## Differential testing

Get which comparisons present enough representation for a two-group comparison
```{r whichComparisons}
cnt <- table(S$orig.ident, Idents(S))
print(cnt)

# Criteria: at least 100 cells in each group
cont <- sapply(setdiff(rownames(cnt), "Healthy"), function(idx) {
			       colSums(rbind(cnt["Healthy",], cnt[idx,])>100)==2
			}, simplify=FALSE)
print(cont)
# Remove unsufficient representation cell types
cont <- lapply(cont, function(z) names(z[z]))
print(cont)
```

Go over each organoid to be compared to Healthy, go over each cell type for the two-group comparison
```{r}
# Create a list of lists to store the result
DGE_res <- setNames(vector("list", length=length(cont)), names(cont))
DGE_res <- sapply(names(DGE_res), function(gr) setNames(vector("list", length=length(cont[[gr]])), cont[[gr]]), simplify=FALSE)
# Loop recursively (2-levels) to get DGE_results
for(gr in names(DGE_res)) {
	for(cell in names(DGE_res[[gr]])) {
		Sx <- S[, S$orig.ident %in% c("Healthy", gr) & Idents(S) == cell]
		Idents(Sx) <- Sx$orig.ident
		dge <- FindMarkers(Sx, ident.1=gr, ident.2="Healthy",
				   test.use="wilcox", only.pos=TRUE)
		dge$gene <- rownames(dge)
 		dge <- dge[, c("gene", colnames(dge)[-ncol(dge)])]
		# store DGE_result
		DGE_res[[gr]][[cell]] <- dge
	}
}
```

Show how many DEGs 
```{r cnt_dges}
lapply(DGE_res, function(resX) lapply(resX, function(z) table(z$p_val_adj < 0.05)))
```


## GO term analysis 

```{r}
GO <- getGmt("../data/Prior/c5.all.v7.0.symbols.gmt")
GO <- geneIds(GO)
GO_t2g <- data.frame(term=unlist(sapply(names(GO), function(k) rep(k, length(GO[[k]])))),
		     gene = unlist(GO))
```

```{r}
GO_res <- lapply(DGE_res, function(k)  {
		ng <- lapply(k, function(z) rownames(z)[z$p_val_adj < 0.05])
 		en <- lapply(ng, function(x) enricher(x, TERM2GENE=GO_t2g, minGSSize = 10, maxGSSize = 200, pvalueCutoff = 0.05))
 		return(en)
	})

GO_res <- lapply(GO_res, function(k) lapply(k, function(z) z@result))
```


## KEGG analysis 

```{r}
KEGG <- getGmt("../data/Prior/c2.cp.kegg.v7.2.symbols.gmt")
KEGG <- geneIds(KEGG)
KEGG_t2g <- data.frame(term=unlist(sapply(names(KEGG), function(k) rep(k, length(KEGG[[k]])))),
		     gene = unlist(KEGG))
```

```{r}
KEGG_res <- lapply(DGE_res, function(k)  {
		ng <- lapply(k, function(z) rownames(z)[z$p_val_adj < 0.05])
 		en <- lapply(ng, function(x) enricher(x, TERM2GENE=KEGG_t2g, minGSSize = 10, maxGSSize = 200, pvalueCutoff = 0.05))
 		return(en)
	})
KEGG_res <- lapply(KEGG_res, function(k) lapply(k, function(z) z@result))
```

## Meta-analysis of DEGs
```{r}
META_res <- setNames(vector("list", length=1), "TPC")
for(cell in names(META_res)) {
	Sx <- S[, Idents(S) == cell]
	Idents(Sx) <- ifelse(Sx$orig.ident=="Healthy", "Healthy", "PDK-KO")

	Si <- Sx[, Sx$orig.ident %in% c("PDK1KO", "Healthy")]
	Si <- RenameCells(Si, paste0(Si$orig.ident,"_",colnames(Si)))
	Si$orig.ident <- "PDK1KO"

	Sj <- Sx[, Sx$orig.ident %in% c("PDK2KO", "Healthy")]
	Sj <- RenameCells(Sj, paste0(Sj$orig.ident,"_",colnames(Sj)))
	Sj$orig.ident <- "PDK2KO"

	Sx <- merge(Si, Sj)

	dge <- FindConservedMarkers(Sx, ident.1="PDK-KO", ident.2="Healthy",
				    grouping.var="orig.ident",
			   test.use="wilcox", only.pos=TRUE)
	dge$gene <- rownames(dge)
	dge <- dge[, c("gene", colnames(dge)[-ncol(dge)])]
	# store META_result
	META_res[[cell]] <- dge
}
```

ORA on these meta-signature
```{r}
KEGG_out<- lapply(META_res, function(k)  {
		ng <- rownames(k)[k$PDK1KO_p_val_adj< 0.05 & k$PDK2KO_p_val_adj < 0.05]
 		en <- enricher(ng, TERM2GENE=KEGG_t2g, minGSSize = 10, maxGSSize = 200, pvalueCutoff = 0.05)
 		return(en)
	})
```

```{r}
GO_out<- lapply(META_res, function(k)  {
		ng <- rownames(k)[k$PDK1KO_p_val_adj< 0.05 & k$PDK2KO_p_val_adj < 0.05]
 		en <- enricher(ng, TERM2GENE=GO_t2g, minGSSize = 10, maxGSSize = 200, pvalueCutoff = 0.05)
 		return(en)
	})
```

```{r}
ADPKD_hallmarks <- getGmt("../data/Prior/ADPKD_sets.gmt")
ADPKD <- geneIds(ADPKD_hallmarks)
ADPKD_t2g <- data.frame(term=unlist(sapply(names(ADPKD), function(k) rep(k, length(ADPKD[[k]])))),
		     gene = unlist(ADPKD))

ADPKD_out<- lapply(META_res, function(k)  {
		ng <- rownames(k)[k$PDK1KO_p_val_adj< 0.05 & k$PDK2KO_p_val_adj < 0.05]
 		en <- enricher(ng, TERM2GENE=ADPKD_t2g, minGSSize = 10, maxGSSize = 200, pvalueCutoff = 0.05)
 		return(en)
	})

```


```{r barplot_ora}
barplot_ora <- function(res, N=10) {
	res$geneID2 <- sapply(res$geneID, function(z) {
				      genes <- strsplit(z, split="/")[[1]]
				      if(length(genes)>10) {
					      txt <- paste0(paste(head(genes, 7), collapse="/"),"...")
				      } else {
					      txt <- z
				      }
				      return(txt)

	})
	res$ID <- factor(res$ID, levels=rev(res$ID))
	res$Ann <- paste0(res$GeneRatio,": ",res$geneID2)

	#NOTE: result is ordered by significance (pval) by default
	gg <- ggplot(head(res, N), aes(y=-log10(p.adjust), x=ID)) + 
		geom_bar(stat="identity", width=0.7, fill="grey") +
		#   geom_point(size=7) +
		#   geom_segment(aes(xend=ID), yend=0) + 
		#   expand_limits(y=0) + 
		geom_text(mapping=aes(y=0.01, x=ID, label=Ann), hjust=0, size=4) + 
		theme_cowplot() +
		geom_hline(yintercept = -log10(0.05), lty=2, alpha=0.5) + 
		coord_flip() +
		ylab("-log10(adjusted P-value)") +
		theme(legend.position="bottom",
		      legend.key.width=unit(1.5,"cm"),
		      legend.title=element_text(size=16),
		      legend.text=element_text(size=16),
		      axis.title.y=element_blank(),
		      axis.text.y=element_text(size=19),
		      axis.text.x=element_text(size=20),
		      axis.title.x=element_text(size=18))
	
	return(gg)

}
```



```{r organoid_KEGG, fig.width=17, fig.height=7, dpi=300}
gg_kegg <- barplot_ora(KEGG_out$TPC@result, 10)
print(gg_kegg)
```

```{r organoid_GO, fig.width=17, fig.height=7, dpi=300}
gg_go <- barplot_ora(GO_out$TPC@result, 10)
print(gg_go)
```

```{r organoid_ADPKDhallmarks, fig.width=17, fig.height=3, dpi=300}
gg_pkd <- barplot_ora(ADPKD_out$TPC@result, 10)
print(gg_pkd)
```


## Save

```{r Excel}
for(gr in names(cont)) {
	## Excel file
	xlsx_file <- paste0(OUTDIR, "/", gr, "_vs_Healthy.xlsx")
	wb <- createWorkbook()

	for(cell in cont[[gr]]) {
		RES <- ls(pattern="_res$")		
		for(RESx in RES) {
			tag <- gsub("_res$", "", RESx)
			tab <- get(RESx)[[gr]][[cell]]
			sheet_name <- paste0(cell,"_",tag)
			addWorksheet(wb, sheetName = sheet_name)
			writeData(wb, sheet = sheet_name, x=tab, rowNames=FALSE)
		}
	}

	## save file 
	saveWorkbook(wb, file =  xlsx_file, overwrite = TRUE)
}
```

```{r Excel2}
## Excel file
xlsx_file <- paste0(OUTDIR, "/meta-analysis_PDK-KO_vs_Healthy.xlsx")
wb <- createWorkbook()
for(cell in names(META_res)) {
	tag <- "DiffExpr_metanalysis"
	tab <- META_res[[cell]]
	sheet_name <- paste0(cell,"_",tag)
	addWorksheet(wb, sheetName = sheet_name)
	writeData(wb, sheet = sheet_name, x=tab, rowNames=FALSE)
}

## save file 
saveWorkbook(wb, file =  xlsx_file, overwrite = TRUE)

```

