---
title: "CK121 sorted CD24+ cells : cell assignment"
author: "Javier Perales-Paton - javier.perales@bioquant.uni-heidelberg.de"
license: "GPL-v3"
output: github_document
---


```{r setup, include=FALSE}
Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio/bin/pandoc")
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and auxiliar functions
```{r libs}
set.seed(1234)
suppressPackageStartupMessages(require(Seurat))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(GSEABase))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(genesorteR))
suppressPackageStartupMessages(require(ComplexHeatmap))
suppressPackageStartupMessages(require(clustree))
suppressPackageStartupMessages(require(cowplot))
source("../src/seurat_fx.R")
```

## Load SeuratObject with initial clustering outcome
```{r load_init}
SeuratObject <- readRDS("./output/1_initial_clustering/data/SeuratObject.rds")
```
## Define output directory
```{r def_out}
# Define output directory
OUTDIR <- paste0("./output/2_cell_assignment/")
if(! dir.exists(OUTDIR)) dir.create(OUTDIR, recursive = TRUE)
```

## Cell marker extraction
We are going to use two methods for the extraction of cell markers.   
1. Differential gene expression using wilcox test from Seurat. Only positive
   and consistent markers will be tested (with default parameters).   
2. GenesorteR to get posterior probabilities for each gene of observing a certain
cell population.

### 1 wilcox test

```{r wilcox, cache=FALSE}
up <- setNames(vector("list",length=length(levels(SeuratObject))), 
               levels(SeuratObject))
for(idx in names(up)) {
  up.idx <- FindMarkers(SeuratObject,ident.1 = idx, 
                        ident.2 = setdiff(levels(SeuratObject), idx), only.pos=T)
  cols_names <- colnames(up.idx)
  
  # Add two extra cols
  up.idx$cluster <- idx
  up.idx$gene <- rownames(up.idx)
  
  up[[idx]] <- up.idx
}
```

### 2 Gene sorter

```{r genesort, cache=FALSE}
sg <- sortGenes(SeuratObject@assays$RNA@data, Idents(SeuratObject))

#define a small set of markers
#mm = getMarkers(sg, quant = 0.975)

#cluster genes and make a heatmap
#pp = plotMarkerHeat(sg$inputMat, sg$inputClass, mm$markers, clusterGenes=TRUE, outs = TRUE)

#pp$gene_class_info #gene clusters

#the top 25 genes for each cluster by specificity scores
top_markers = apply(sg$specScore, 2, function(x) names(head(sort(x, decreasing = TRUE), n = 25)))

```

Finally we save the markers for manual exploration:

```{r init_markers_save}
MARKERS_OUTDIR <- paste0(OUTDIR,"/Markers")
if(! dir.exists(MARKERS_OUTDIR)) dir.create(MARKERS_OUTDIR, recursive = TRUE)

# Wilcox
for(idx in names(up)) {
  write.table(up[[idx]][,c("cluster", "gene", cols_names)],
              file = paste0(MARKERS_OUTDIR,"/cluster",idx,".tsv"),
              sep="\t",col.names = TRUE, row.names = FALSE, quote=FALSE
  )
}
saveRDS(up, file=paste0(MARKERS_OUTDIR,"/wilcox_up.rds"))

# Genesorter
saveRDS(sg, file=paste0(MARKERS_OUTDIR,"/genesorter_out.rds"))
write.table(as.matrix(sg$specScore), paste0(MARKERS_OUTDIR,"/specScore.tsv"),
            sep="\t",row.names = TRUE,col.names = NA,quote=FALSE)

write.table(as.matrix(sg$condGeneProb), paste0(MARKERS_OUTDIR,"/condGeneProb.tsv"),
            sep="\t",row.names = TRUE,col.names = NA,quote=FALSE)
```


## Cell assignment
We took a look at the markers from the different clusters from previous section. These can be found in the `r MARKERS_OUTDIR` directory.

Looking into them, we could identify different cell lineages that are actually 
contaminant and not expected in the CD24+ sorted cells:
- Hematopoietic cells (PTPRC+) : Many of the initial clusters.
- Endothelial cells (PECAM1+) : One cluster


| Cluster 0 | exclude hematopoietic |
| Cluster 1 | exclude hematopoietic |
| Cluster 2 | epithelium (43% Pax8), SLC12a1 49%, UMOD 13%, LOH, MTAL cell |
| Cluster 3 | exclude hematopoietic |
| Cluster 4 | epithelium 43% AKAP12, 11% SLC34a1, 15% SLC22a6, proximal tubule injured/scattered (STCs) |
| Cluster 5 | 65% AKAP12, SLC22a2+, 22% CLDN1+  STCs |
| Cluster 6 | exclude hematopoietic |
| Cluster 7 | exclude endothelium |
| Cluster 8 | epithelium,  SLC4a1 28%,25% CLCNKA, Type A IC |
| Cluster 9 | epithelium 61% cldn1, 54% AKAP12, 61% KLK6,   PECs also VCAM1 highest cluster PECs |
| cluster 10 | exclude hematopoietic |

```{r reports}
reporters <- list("LOH, MTAL"=c("PAX8", "SLC12A1","UMOD"),
                  "Endothelial"=c("PECAM1"),
                  "TPC"=c("AKAP12", "SLC34A1", "SLC22A6","SLC22A2",
                          "CDH6", "WNT7A", "CLDN7"),
                  "IC-A"=c("SLC4A1","CLCNKA"),
                  "PEC"=c("CLDN1", "AKAP12","KLF6"),
                  "DCT"=c("SLC12A3", "SLC14A2", "AKAP12",
                          "COL4A3", "SCN2A", "AGR3", "DEFB1", "SLC14A2"),
                  "Prolif."=c("MKI67", "CENPF", "CDKN3", "BIRC5", "CDC20",
                                     "NUSAP1", "AKAP12"),
                  "Hematopoietic"=c("PTPRC"))
```

```{r cond_heatmap}
# condGeneProb from genesorteR
Heatmap(as.matrix(sg$condGeneProb)[unlist(reporters),], name="condGeneProb",
        col=c("white","red"),row_names_gp = gpar(fontsize=10),
        cluster_rows = FALSE, cluster_columns = FALSE,
        split = factor(unlist(sapply(names(reporters), function(z) rep(z,length(reporters[[z]])))),
                       levels=names(reporters)),
        row_title_rot = 0,row_gap = unit(0, "mm"),border=TRUE
        )
```

```{r expr_heatmap}
DoHeatmap2(SeuratObject, 
           GS=GeneSetCollection(sapply(names(reporters), 
                                       function(gs) GeneSet(unique(reporters[[gs]]), 
                                                            setName=gs))),
           assay = "RNA", res = 0.5
           )
```

```{r rename_idents}
ren_id <- c("0"="Exclude_hematopoietic",
            "1"="Exclude_hematopoietic",
            "2"="LOH, MTAL cell",
            "3"="Exclude_hematopoietic",
            "4"="TPC_2",
            "5"="TPC_1",
            "6"="Exclude_hematopoietic",
            "7"="Exclude_endothelial",
            "8"="IC-A",
            "9"="PEC",
            "10"="Exclude_hematopoietic")
```



```{r cell_ann}
SeuratObject <- RenameIdents(SeuratObject, ren_id)
SeuratObject$init_assign <- Idents(SeuratObject)
```


## UMAP plot
```{r}
DimPlot(SeuratObject, reduction = "umap", label=TRUE)
```

## Archive processed data for downstream analysis

```{r assign_record}
# new idents
write.table(data.frame("Ident"=SeuratObject@active.ident),
            file=paste0(OUTDIR,"/active_idents.tsv"),
            sep="\t", col.names = NA, row.names = TRUE, quote=TRUE)
```

```{r, mkdir_data}
DATA_DIR <- paste0(OUTDIR,"/data")
if(!dir.exists(DATA_DIR)) dir.create(DATA_DIR)
```

```{r save_S}
saveRDS(SeuratObject, paste0(DATA_DIR,"/SeuratObject.rds"))
```                                                                                                                                                                                                         

## Session info

```{r}
sessionInfo()
{                                                                                                                                                                                                           
sink(file=paste0(OUTDIR,"/sessionInfo.txt"))
print(sessionInfo())
sink()
}
```
