---
title: "PKD1-KO organoid derived from human CD24+ cells: cell assignment"
author: "Javier Perales-Patón - javier.perales@bioquant.uni-heidelberg.de - ORCID: 0000-0003-0780-6683" 
license: "GPL-v3"
output: github_document
---


```{r setup, include=FALSE}
Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio/bin/pandoc")
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and auxiliar functions
```{r libs}
set.seed(1234)
suppressPackageStartupMessages(require(Seurat))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(GSEABase))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(genesorteR))
suppressPackageStartupMessages(require(ComplexHeatmap))
suppressPackageStartupMessages(require(clustree))
suppressPackageStartupMessages(require(cowplot))
suppressPackageStartupMessages(require(openxlsx))
source("../src/seurat_fx.R")
```

## Load SeuratObject with initial clustering outcome
```{r load_init}
SeuratObject <- readRDS("./output/1_initial_clustering/data/SeuratObject.rds")
```
## Define output directory
```{r def_out}
# Define output directory
OUTDIR <- paste0("./output/2_cell_assignment/")
if(! dir.exists(OUTDIR)) dir.create(OUTDIR, recursive = TRUE)
```

## Cell marker extraction
We are going to use two methods for the extraction of cell markers.   
1. Differential gene expression using wilcox test from Seurat. Only positive
   and consistent markers will be tested (with default parameters).   
2. GenesorteR to get posterior probabilities for each gene of observing a certain
cell population.

### 1 wilcox test

```{r wilcox, cache=FALSE}
up <- setNames(vector("list",length=length(levels(SeuratObject))), 
                levels(SeuratObject))
for(idx in names(up)) {
  up.idx <- FindMarkers(SeuratObject,ident.1 = idx, 
                        ident.2 = setdiff(levels(SeuratObject), idx), only.pos=T)
  cols_names <- colnames(up.idx)
   
  # Add two extra cols
  up.idx$cluster <- idx
  up.idx$gene <- rownames(up.idx)
   
  up[[idx]] <- up.idx
}
```

### 2 Gene sorter

```{r genesort, cache=FALSE}
sg <- sortGenes(SeuratObject@assays$RNA@data, Idents(SeuratObject))
```

Finally we save the markers for manual exploration:
```{r init_markers_save, warning=FALSE}
saveMarkers.Excel(up, sg, SeuratObject, OUTDIR)
saveMarkers.CSV(up, OUTDIR)
saveMarkers.CSV(sg, OUTDIR)
```

## Cell assignment
We took a look at the markers from the different clusters from previous section. 
These can be found in the `r OUTDIR` directory.

We took a look at the markes from previous section, making the following summary


```{r cellQual}
FeaturePlot(SeuratObject, 
	    features=c("percent.mt", "nFeature_RNA", 
		       "S.Score", "G2M.Score"), 
	    label=TRUE, label.size=7)
```

```{r CellCycle, fig.width=7, fig.height=5}
VlnPlot(SeuratObject, features=c("S.Score", "G2M.Score"))
```


```{r rename_idents}
ren_id <- c("0"="DCT-like_1",
            "1"="TPC_1:LowCov",
            "2"="TPC_2",
            "3"="DCT-like_2")
```



```{r cell_ann}
stopifnot(all(names(ren_id) %in% levels(SeuratObject)))
stopifnot(length(ren_id) == length(levels(SeuratObject)))

SeuratObject <- RenameIdents(SeuratObject, ren_id)
SeuratObject$init_assign <- factor(as.character(Idents(SeuratObject)))

colnames(sg$condGeneProb) <- ren_id[colnames(sg$condGeneProb)]
colnames(sg$specScore) <- ren_id[colnames(sg$specScore)]
```

```{r load_markers}
AdultOrganoid <- getGmt("../data/Prior/adultorganoid_markers.gmt")
AdultOrganoid <- GeneSetCollection(lapply(AdultOrganoid, function(GS) {
			  GS@geneIds <- intersect(GS@geneIds, rownames(SeuratObject))
 			  GS@geneIds <- intersect(GS@geneIds, rownames(sg$condGeneProb))
			  return(GS)
	    }))
AdultOrganoid <- AdultOrganoid[unlist(lapply(geneIds(AdultOrganoid), length)) > 0]
```

```{r AdultOrganoid_mhp, fig.width=16, fig.height=14, dpi=300, warning=FALSE}
DoMultiHeatmap(SeuratObject, sg, GSC = AdultOrganoid, assay = "RNA", show_plot = TRUE)
```

```{r AdultOrganoid_dplot, fig.width=16, fig.height=14, dpi=300, warning=FALSE}
DoMultiDotPlot(SeuratObject, sg, GSC = AdultOrganoid)
```

```{r ClustReport, fig.width=16, fig.height=17, dpi=300, warning=FALSE}
DoClustReport(SeuratObject, sg, AdultOrganoid, show_NCells=TRUE)
```

## UMAP plot
```{r}
DimPlot(SeuratObject, reduction = "umap", label=TRUE) +
	coord_cartesian(clip = "off")
```

## Archive processed data for downstream analysis

```{r assign_record}
# new idents
write.table(data.frame("Ident"=SeuratObject@active.ident),
            file=paste0(OUTDIR,"/active_idents.tsv"),
            sep="\t", col.names = NA, row.names = TRUE, quote=TRUE)
```

```{r, mkdir_data}
DATA_DIR <- paste0(OUTDIR,"/data")
if(!dir.exists(DATA_DIR)) dir.create(DATA_DIR)
```

```{r save_S}
saveRDS(SeuratObject, paste0(DATA_DIR,"/SeuratObject.rds"))
saveRDS(up, file=paste0(OUTDIR,"/wilcox_up.rds"))
saveRDS(sg, file=paste0(OUTDIR,"/genesorter_out.rds"))

# After Rename Idents
saveMarkers.Excel(up, sg, SeuratObject, OUTDIR)
```                                                                                                                                                                                                         

## Session info

```{r}
sessionInfo()
{                                                                                                                                                                                                           
sink(file=paste0(OUTDIR,"/sessionInfo.txt"))
print(sessionInfo())
sink()
}
```
