---
title: "CK5 early organoid : cell assignment"
author: "Javier Perales-Paton - javier.perales@bioquant.uni-heidelberg.de"
license: "GPL-v3"
output: github_document
---


```{r setup, include=FALSE}
Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio/bin/pandoc")
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and auxiliar functions
```{r libs}
set.seed(1234)
suppressPackageStartupMessages(require(Seurat))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(GSEABase))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(genesorteR))
suppressPackageStartupMessages(require(ComplexHeatmap))
suppressPackageStartupMessages(require(clustree))
suppressPackageStartupMessages(require(cowplot))
source("../src/seurat_fx.R")
```

## Load SeuratObject with initial clustering outcome
```{r load_init}
SeuratObject <- readRDS("./output/1_initial_clustering/data/SeuratObject.rds")
```
## Define output directory
```{r def_out}
# Define output directory
OUTDIR <- paste0("./output/2_cell_assignment/")
if(! dir.exists(OUTDIR)) dir.create(OUTDIR, recursive = TRUE)
```

## Cell marker extraction
We are going to use two methods for the extraction of cell markers.   
1. Differential gene expression using wilcox test from Seurat. Only positive
   and consistent markers will be tested (with default parameters).   
2. GenesorteR to get posterior probabilities for each gene of observing a certain
cell population.

### 1 wilcox test

```{r wilcox, cache=FALSE}
up <- setNames(vector("list",length=length(levels(SeuratObject))), 
               levels(SeuratObject))
for(idx in names(up)) {
  up.idx <- FindMarkers(SeuratObject,ident.1 = idx, 
                        ident.2 = setdiff(levels(SeuratObject), idx), only.pos=T)
  cols_names <- colnames(up.idx)
  
  # Add two extra cols
  up.idx$cluster <- idx
  up.idx$gene <- rownames(up.idx)
  
  up[[idx]] <- up.idx
}
```

### 2 Gene sorter

```{r genesort, cache=FALSE}
sg <- sortGenes(SeuratObject@assays$RNA@data, Idents(SeuratObject))

#define a small set of markers
#mm = getMarkers(sg, quant = 0.975)

#cluster genes and make a heatmap
#pp = plotMarkerHeat(sg$inputMat, sg$inputClass, mm$markers, clusterGenes=TRUE, outs = TRUE)

#pp$gene_class_info #gene clusters

#the top 25 genes for each cluster by specificity scores
top_markers = apply(sg$specScore, 2, function(x) names(head(sort(x, decreasing = TRUE), n = 25)))

```

Finally we save the markers for manual exploration:

```{r init_markers_save}
MARKERS_OUTDIR <- paste0(OUTDIR,"/Markers")
if(! dir.exists(MARKERS_OUTDIR)) dir.create(MARKERS_OUTDIR, recursive = TRUE)

# Wilcox
for(idx in names(up)) {
  write.table(up[[idx]][,c("cluster", "gene", cols_names)],
              file = paste0(MARKERS_OUTDIR,"/cluster",idx,".tsv"),
              sep="\t",col.names = TRUE, row.names = FALSE, quote=FALSE
  )
}
saveRDS(up, file=paste0(MARKERS_OUTDIR,"/wilcox_up.rds"))

# Genesorter
saveRDS(sg, file=paste0(MARKERS_OUTDIR,"/genesorter_out.rds"))
write.table(as.matrix(sg$specScore), paste0(MARKERS_OUTDIR,"/specScore.tsv"),
            sep="\t",row.names = TRUE,col.names = NA,quote=FALSE)

write.table(as.matrix(sg$condGeneProb), paste0(MARKERS_OUTDIR,"/condGeneProb.tsv"),
            sep="\t",row.names = TRUE,col.names = NA,quote=FALSE)
```


## Cell assignment
We took a look at the markers from the different clusters from previous section. These can be found in the `r MARKERS_OUTDIR` directory.

We took a look at the markes from previous section, making the following summary

| Cluster | rational |
|:---:|:---:|
| Cluster 0 | SLC34a2 high, SERPINE2+, EMP3+ proximal tubule  AKAP12 25% PT like |
| Cluster 1 | 35% SLC12a3, SLC14a2 5%, Distal Convoluted Tubule (DCT) AKAP12 25% |
| Cluster 2 | AKAP12 68%, CDH6+, WNT7a+, CLDN7+, AKAP12 68% TPCs |
| Cluster 3 | 3% SLC12a1, SLC34a2 high, NCOA7+, LTF+ AKAP12 15% PT like |
| Cluster 4 | NEAT1+, XIST+, MALAT1+, AKAP12 16% delete mito genes, looks odd |
| Cluster 5 | 25% SLC12a3 Distal Convoluted Tubule,Col4a3+, Col4a4+, SCN2a+ AGR3+, DEFB1+,  SLC14a2 3%, Distal Convoluted Tubule (DCT) no AKAP12 |
| Cluster 6 | MKI67+, , CENPF, CDKN3, BIRC5, CDC20, NUSAP1, AKAP12 30% proliferating PT-like |
| Cluster 7 | 37% AQP1+, VCAM1 59%, AKAP12 60%, THBS2+, NUPR1+ PEC like |
| Cluster 8 | no Pax8, CD133 (PROM1) 13%, no high spec score, ALDH1A3, AKAP12 11% (exclude high ribosomal probably garbadge) |
| Cluster 9 | CD133 0%, VCAM1 33%, THY1+, KRT23, SLIT3, CD13 (ANPEP), UPK3B, AKAP12 28%, EPCAM +PEC like |

```{r reports}
reporters <- list("PT-like"=c("SLC12A1", "SLC12A2", "NCOA7", "LTF",
                              "SLC34A2", "SERPINE2", "EMP3", "AKAP12"),
                  "DCT"=c("SLC12A3", "SLC14A2", "AKAP12",
                          "COL4A3", "SCN2A", "AGR3", "DEFB1", "SLC14A2"),
                  "TPC"=c("AKAP12", "CDH6", "WNT7A", "CLDN7"),
                  "Prolif.PT-like"=c("MKI67", "CENPF", "CDKN3", "BIRC5", "CDC20",
                                     "NUSAP1", "AKAP12"),
                  "PEC-like"=c("AQP1", "VCAM1", "AKAP12", "THBS2", "NUPR1",
                               "THY1", "KRT23","SLIT3", "ANPEP", "UPK3B"),
                  "Others"=c("PAX8", "PROM1"))
```

```{r cond_heatmap}
# condGeneProb from genesorteR
Heatmap(as.matrix(sg$condGeneProb)[unlist(reporters),], name="condGeneProb",
        col=c("white","red"),row_names_gp = gpar(fontsize=10),
        cluster_rows = FALSE, cluster_columns = FALSE,
        split = factor(unlist(sapply(names(reporters), function(z) rep(z,length(reporters[[z]])))),
                       levels=names(reporters)),
        row_title_rot = 0,row_gap = unit(0, "mm"),border=TRUE
        )
```

```{r expr_heatmap}
DoHeatmap2(SeuratObject, 
           GS=GeneSetCollection(sapply(names(reporters), 
                                       function(gs) GeneSet(unique(reporters[[gs]]), 
                                                            setName=gs))),
           assay = "RNA", res = 0.5
           )
```

Expression of mitochondrial and ribosomal genes are going to discard cluster 4 and 8, 
respectively. Moreover, Cluster 8 has no PAX8 expression (kidney marker), resulting in 
an aberrant cell population unexpected in this context (contaminant).

```{r excluding}
DoHeatmap2(SeuratObject, 
           GS=GeneSetCollection(GeneSet(grep("^RPL",rownames(SeuratObject), value=TRUE), 
                                        setName="Ribosomal"),
                                GeneSet(grep("^MT-", rownames(SeuratObject), value=TRUE), 
                                        setName="Mitochondrial")),
           assay = "RNA", res = 0.5
           )
```

```{r rename_idents}
ren_id <- c("0"="PT-like_1",
            "1"="DCT_1",
            "2"="TPC",
            "3"="PT-like_2",
            "4"="Exclude_highmito",
            "5"="DCT_2",
            "6"="Proliferating_PT-like",
            "7"="PEC-like_1",
            "8"="Exclude_highribosomal",
            "9"="PEC-like_2")
```



```{r cell_ann}
SeuratObject <- RenameIdents(SeuratObject, ren_id)
SeuratObject$init_assign <- Idents(SeuratObject)
```


## UMAP plot
```{r}
DimPlot(SeuratObject, reduction = "umap", label=TRUE)
```

## Archive processed data for downstream analysis

```{r assign_record}
# new idents
write.table(data.frame("Ident"=SeuratObject@active.ident),
            file=paste0(OUTDIR,"/active_idents.tsv"),
            sep="\t", col.names = NA, row.names = TRUE, quote=TRUE)
```

```{r, mkdir_data}
DATA_DIR <- paste0(OUTDIR,"/data")
if(!dir.exists(DATA_DIR)) dir.create(DATA_DIR)
```

```{r save_S}
saveRDS(SeuratObject, paste0(DATA_DIR,"/SeuratObject.rds"))
```                                                                                                                                                                                                         

## Session info

```{r}
sessionInfo()
{                                                                                                                                                                                                           
sink(file=paste0(OUTDIR,"/sessionInfo.txt"))
print(sessionInfo())
sink()
}
```
