---
title: "CK5 early organoid : Refine cell clustering"
author: "Javier Perales-Paton - javier.perales@bioquant.uni-heidelberg.de"
license: "GPL-v3"
output: github_document
---
  
```{r setup, include=FALSE}
Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio/bin/pandoc")
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and auxiliar functions
```{r libs}
set.seed(1234)
suppressPackageStartupMessages(require(Seurat))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(GSEABase))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(genesorteR))
suppressPackageStartupMessages(require(ComplexHeatmap))
suppressPackageStartupMessages(require(clustree))
suppressPackageStartupMessages(require(cowplot))
source("../src/seurat_fx.R")
```

## Load SeuratObject with initial clustering outcome
```{r load_init}
SeuratObject <- readRDS("./output/2_cell_assignment/data/SeuratObject.rds")
```
## Define output directory
```{r def_out}
# Define output directory
OUTDIR <- paste0("./output/3_refine_clustering/")
if(! dir.exists(OUTDIR)) dir.create(OUTDIR, recursive = TRUE)
```

## Re-clustering and high dim. reduction after filtering contaminant populations

To trace back previous cell boundaries, we store the outcome
```{r init_cl}
if(!"init_seurat_clusters" %in% colnames(SeuratObject@meta.data))
  SeuratObject$init_seurat_clusters <- SeuratObject$seurat_clusters

if(!"init_assign" %in% colnames(SeuratObject@meta.data))
  SeuratObject$init_assign <- Idents(SeuratObject)
```

We remove two cell populations that present with over-expression of genes related to bad quality cells.
We observed that cluster 5 over-express high mitochondrial genes, whereas cluster 9 over-express ribosomal genes.

```{r assign_filter}
# Remove original reference to clustering
SeuratObject <- SeuratObject[, !grepl(":(High|Low)", 
                                      Idents(SeuratObject))]
```

We proceed with the standard pipeline for cell clustering, over-written previous
outcome

```{r re_clust, cache=FALSE, fig.width=11, fig.height=4}
## Feature selection
SeuratObject <- FindVariableFeatures(SeuratObject, selection.method = "vst", nfeatures = 2000)

## PCA 
SeuratObject <- RunPCA(SeuratObject, features = VariableFeatures(SeuratObject), npcs = 50)
print(ElbowPlot(SeuratObject,ndims = 50) + geom_vline(xintercept = 25, col="red"))

## Cell clustering
SeuratObject <- FindNeighbors(SeuratObject, dims = 1:25)
SeuratObject <- FindClusters(SeuratObject, resolution = 0.5)

## Agreement with previous clustering
table("initial"=SeuratObject$init_seurat_clusters,
      "final"=SeuratObject$seurat_clusters)

table("Assigned"=SeuratObject$init_assign,
      "final"=SeuratObject$seurat_clusters)


## UMAP
SeuratObject <- RunUMAP(SeuratObject, dims = 1:25)
d1 <- DimPlot(SeuratObject, group.by = "init_assign") + ggtitle("Initial Cell assignment")
d2 <- DimPlot(SeuratObject) + ggtitle("Re-classification")

print(CombinePlots(list(d1,d2)))
```

## Diagnostics of unsupervised clustering

We will investigate how the selection of multiple resolutions affects the partition
into individual cell clusters.
```{r clustree, fig.width=7, fig.height=10}
clustree(SeuratObject, prefix = "RNA_snn_res.")
```

We also checked whether mithochondrial expression drives the clustering
```{r clustree_mt, fig.width=7, fig.height=10}
clustree(SeuratObject, prefix = "RNA_snn_res.", 
	 node_colour="percent.mt", node_colour_aggr="mean")
```

```{r clustree_nFeat, fig.width=7, fig.height=10}
clustree(SeuratObject, prefix = "RNA_snn_res.", 
	 node_colour="nFeature_RNA", node_colour_aggr="median")
```

```{r vln_badQual_mtNfeat}
VlnPlot(SeuratObject, 
	feature=c("percent.mt", "nFeature_RNA", 
		"Dissociation", "G2M.Score"), 
	pt.size=0.4,
	ncol=2)
```



## Archive processed data for downstream analysis
```{r, mkdir_data}
DATA_DIR <- paste0(OUTDIR,"/data")
if(!dir.exists(DATA_DIR)) dir.create(DATA_DIR)
```

```{r clustering_record}
# 1 Clustering outcome
saveClusteringOutcome(SeuratObject , assay="RNA", fl=paste0(OUTDIR,"/init_clustering.tsv"))
# 2 Initial idents (same as seurat_clusters)
saveActiveIdents(SeuratObject, fl=paste0(OUTDIR,"/active_idents.tsv"))
```

```{r save_S}
saveRDS(SeuratObject, paste0(DATA_DIR,"/SeuratObject.rds"))
```                                                                                                                                                                                                         

## Session info

```{r}
sessionInfo()
{                                                                                                                                                                                                           
sink(file=paste0(OUTDIR,"/sessionInfo.txt"))
print(sessionInfo())
sink()
}
```
