---
title: "CK5 early organoid : final cell assignment" 
author: "Javier Perales-Paton - javier.perales@bioquant.uni-heidelberg.de"
license: "GPL-v3"
output: github_document
---
  
```{r setup, include=FALSE}
Sys.setenv(RSTUDIO_PANDOC="/usr/lib/rstudio/bin/pandoc")
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and auxiliar functions
```{r libs}
set.seed(1234)
suppressPackageStartupMessages(require(Seurat))
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(GSEABase))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(genesorteR))
suppressPackageStartupMessages(require(ComplexHeatmap))
suppressPackageStartupMessages(require(clustree))
suppressPackageStartupMessages(require(cowplot))
source("../src/seurat_fx.R")
```

## Load SeuratObject with initial clustering outcome
```{r load_init}
SeuratObject <- readRDS("./output/3_refine_clustering//data/SeuratObject.rds")
```
## Define output directory
```{r def_out}
# Define output directory
OUTDIR <- paste0("./output/4_final_assignment/")
if(! dir.exists(OUTDIR)) dir.create(OUTDIR, recursive = TRUE)
```


## Identify the proliferating PT-like cell population
```{r cell_phase}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

SeuratObject <- CellCycleScoring(SeuratObject, s.features = s.genes, g2m.features = g2m.genes, set.ident = FALSE)
d3 <- DimPlot(SeuratObject, group.by = "Phase") + ggtitle("Cell Cycle Phase")
d4 <- DimPlot(SeuratObject) + ggtitle("Re-classification")

print(CombinePlots(list(d3,d4)))

```


## Cell marker extraction
We are going to use two methods for the extraction of cell markers.   
1. Differential gene expression using wilcox test from Seurat. Only positive
   and consistent markers will be tested (with default parameters).   
2. GenesorteR to get posterior probabilities for each gene of observing a certain
cell population.

### 1 wilcox test

```{r wilcox, cache=FALSE}
up <- setNames(vector("list",length=length(levels(SeuratObject))), 
               levels(SeuratObject))
for(idx in names(up)) {
  up.idx <- FindMarkers(SeuratObject,ident.1 = idx, 
                        ident.2 = setdiff(levels(SeuratObject), idx), only.pos=T)
  cols_names <- colnames(up.idx)
  
  # Add two extra cols
  up.idx$cluster <- idx
  up.idx$gene <- rownames(up.idx)
  
  up[[idx]] <- up.idx
}
```

### 2 Gene sorter

```{r genesort, cache=FALSE}
sg <- sortGenes(SeuratObject@assays$RNA@data, Idents(SeuratObject))

#define a small set of markers
#mm = getMarkers(sg, quant = 0.975)

#cluster genes and make a heatmap
#pp = plotMarkerHeat(sg$inputMat, sg$inputClass, mm$markers, clusterGenes=TRUE, outs = TRUE)

#pp$gene_class_info #gene clusters

#the top 25 genes for each cluster by specificity scores
top_markers = apply(sg$specScore, 2, function(x) names(head(sort(x, decreasing = TRUE), n = 25)))

```

Finally we save the markers for manual exploration:

```{r init_markers_save}
MARKERS_OUTDIR <- paste0(OUTDIR,"/Markers")
if(! dir.exists(MARKERS_OUTDIR)) dir.create(MARKERS_OUTDIR, recursive = TRUE)

# Wilcox
## TSV
write.table(do.call("rbind",up)[c("cluster", "gene", cols_names)],
	    file = paste0(MARKERS_OUTDIR,"/",Project(SeuratObject),"_wilcox_DEGs_up.csv"),
	    sep=",", col.names=TRUE, row.names = FALSE, quote=FALSE
	    )
## bin
saveRDS(up, file=paste0(MARKERS_OUTDIR,"/wilcox_up.rds"))

# Genesorter
saveRDS(sg, file=paste0(MARKERS_OUTDIR,"/genesorter_out.rds"))
write.table(as.matrix(sg$specScore), 
	    paste0(MARKERS_OUTDIR,"/",Project(SeuratObject),"_specScore.csv"),
            sep=",",row.names = TRUE,col.names = NA,quote=FALSE)

write.table(as.matrix(sg$condGeneProb), 
	    paste0(MARKERS_OUTDIR,"/",Project(SeuratObject),"_condGeneProb.csv"),
            sep=",",row.names = TRUE,col.names = NA,quote=FALSE)
```

## Final cell assignment
The final table of markers on the second round of cell clustering give the following results:

| Cluster | rational |
|:---|:---|
| Cluster 0 | TPC |
| Cluster 1 | TPC |
| Cluster 2 | DCT (GATA3, SLC12A3, KRT19) |
| Cluster 3 | DCT (SPP1, SLC12A3) |
| Cluster 4 | TPC |
| Cluster 5 | Proliferating TPC |
| Cluster 6 | PEC-like |
| Cluster 7 | PEC-like |

```{r cell_assign2}
ren_id <- c("0"="TPC_1",
            "1"="TPC_2",
            "2"="DCT_1",
            "3"="DCT_2",
            "4"="TPC_3",
            "5"="Prolif.TPC",
            "6"="PEC-like_1",
            "7"="PEC-like_2")
SeuratObject <- RenameIdents(SeuratObject, ren_id)
```

### TSNE
```{r tsne_final}
DimPlot(SeuratObject, reduction="tsne")
```

### UMAP
```{r umap_final}
DimPlot(SeuratObject, reduction="umap")
```


## Archive processed data for downstream analysis

```{r assign_outcome}
# final idents
write.table(data.frame("Ident"=SeuratObject@active.ident,
		       "seurat_clusters"=SeuratObject$seurat_clusters),
            file=paste0(OUTDIR,"/active_idents.csv"),
            sep=",", col.names = NA, row.names = TRUE, quote=TRUE)
```

```{r, mkdir_data}
DATA_DIR <- paste0(OUTDIR,"/data")
if(!dir.exists(DATA_DIR)) dir.create(DATA_DIR)
```

```{r save_S}
saveRDS(SeuratObject, paste0(DATA_DIR,"/SeuratObject.rds"))
```                                                                                                                                                                                                         

## Session info

```{r}
sessionInfo()
{                                                                                                                                                                                                           
sink(file=paste0(OUTDIR,"/sessionInfo.txt"))
print(sessionInfo())
sink()
}
```
